Global study of well-being and internet adoption | Online analysis supplement

### 2021-11-30 03-model-estimate.Rmd
Refactor fitting: map() over models and outcomes

# MCMC settings
options(brms.backend = "cmdstanr")
hmc <- list(
  chains = 4,
  cores = 4,
  threads = 3,
  iter = 2000, 
  warmup = 1000,
  refresh = 50,
  adapt_delta = .90, 
  max_treedepth = 10
)
```

@@

- Transform edge values of internet variable to fit beta
- center year on 2010 (roughly in the middle)
- Intercept and simple effects at 2010
- 1 unit: year
- create 1-year lagged internet adoption variables
- Intercepts are country/grand means
- 1 unit: 100%
- Contrast code factors

```{r}
# Transform edge proportions to fit beta distribution
# when this is used as an outcome
itu <- itu %>% 
  mutate(ib = if_else(internet == 0, .0000001, internet)) %>% 
  mutate(ib = if_else(internet == 1, .9999999, internet))

@@

```{r}
# Give each outcome value a tag to ensure it is used in MV model
dat <- dat %>% 
  mutate(y = !is.na(val))

# Indicate unique ITU values for MV model
dat <- dat %>% 
  # Unique values exist for each country-year
  group_by(region, country, year) %>% 
  mutate(itu = 1:n() == 1) %>% 
  ungroup()
```

@@

## Model 1

@@

### Estimate

@@

bf_1_se <- bf(
  bf(val | resp_se(se, sigma = TRUE) ~ 
       year * sex + year * age +
       (year * sex + year * age |c| country) +
       (year * sex + year * age |r| region),
     sigma ~ 1
  ),
  family = gaussian(),
  unused = ~ cause
)
```

```{r}
fits_1 <- dat %>%
  group_by(cause) %>% 
  summarise(
    fit = list(
      brm(
        formula = bf_itu + bf_1 + set_rescor(FALSE), 
        data = cur_data_all(),
        chains = hmc$chains, 
        cores = hmc$cores, 
        threads = hmc$threads,
        iter = hmc$iter, 
        warmup = hmc$warmup, 
        refresh = hmc$refresh,
        control = list(
          adapt_delta = hmc$adapt_delta, 
          max_treedepth = hmc$max_treedepth
        ),
        file = str_glue("output/brm-1-{cur_group()}")
      )
    )
  )

# fits_1_se <- dat %>%
#   filter(!(cause %in% c("Anxiety", "Depression", "Selfharm")))
#   group_by(cause) %>%
#   summarise(
#     fit_se = list(
#       brm(
#         formula = bf_itu + bf_1_se + set_rescor(FALSE),
#         data = cur_data_all(),
#         chains = hmc$chains,
#         cores = hmc$cores,
#         threads = hmc$threads,
#         iter = hmc$iter,
#         warmup = hmc$warmup,
#         refresh = hmc$refresh,
#         control = list(
#           adapt_delta = hmc$adapt_delta,
#           max_treedepth = hmc$max_treedepth
#         ),
#         file = str_glue("output/brm-1-se-{cur_group()}")
#       )
#     )
#   )
```

@@

```{r}
f1d <- fits_1 %>% 
  mutate(
    out = 
      map(
        fit, 
        ~as_draws_df(., variable = "^b_|^sd_|^cor_", regex = TRUE) %>% 
          summarise_draws(samples = length, default_convergence_measures())
      )
  ) %>% 
  select(-fit) %>% 
  unnest(out)
f1d %>%  
  kbl(
    caption = "HMC Diagnostics, Model 1c",
    digits = c(0,0,0,2,0,0)
  ) %>% 
  kable_paper(html_font = Font) %>% 
  column_spec(
    4, 
    color = "white",
    background = spec_color(
      f1d$rhat, 
      direction = 1, 
      option = "C", 
      end = .7
    )
  )
```

```{r}
tmp <- fits_1 %>% 
  mutate(
    out = map(
      fit, 
      ~pp_check(
        .x, 
        ndraws = 10, 
        type = "dens_overlay_grouped",
        group = "region",
        resp = "val", 
        newdata = .x$data
      ) + 
        facet_wrap("group", nrow = 1, scales = "free")
    )
  )
wrap_plots(tmp$out, nrow = 3) &
  scale_x_continuous(
    breaks = pretty_breaks(2),
    labels = function(x) percent(x/100)
  )
```

## Model 2

Model 2 is a modification of Model 1. First, we now switch to looking at internet use more specifically as a predictor, and therefore implement a univariate model of each outcome. Second, we include within- and between-country centred 1-year-lagged internet adoption values as predictors in the model.

We fit this model in two increasingly complex forms, corresponding to models 1b and 1c, but with the addition of the internet adoption predictors, and one outcome model only.

2c takes about 5 hours per 1000 iterations.

```{r}
bf_2 <- bf(
  val ~ 
    year * sex + year * age + i1_cw * sex + i1_cw * age + i1_cb + 
    (year * sex + year * age + i1_cw * sex + i1_cw * age | country) +
    (year * sex + year * age + i1_cw * sex + i1_cw * age | region),
  family = gaussian(),
  unused = ~ cause
)

# Compile model separately for each outcome to give independent default priors
fits_2 <- dat %>%
  group_by(cause) %>% 
  summarise(
    fit = list(
      brm(
        formula = bf_2, 
        data = cur_data_all(),
        chains = hmc$chains, 
        cores = hmc$cores, 
        threads = hmc$threads,
        iter = hmc$iter, 
        warmup = hmc$warmup, 
        refresh = hmc$refresh,
        control = list(
          adapt_delta = hmc$adapt_delta, 
          max_treedepth = hmc$max_treedepth
        ),
        file = str_glue("output/brm-2-{cur_group()}")
      )
    )
  )

# fits_2 <- dat %>%
#   group_by(cause) %>% 
#   summarise(
#     fit_se = list(
#       brm(
#         formula = bf_2_se, 
#         data = cur_data_all(),
#         chains = hmc$chains, 
#         cores = hmc$cores, 
#         threads = hmc$threads,
#         iter = hmc$iter, 
#         warmup = hmc$warmup, 
#         refresh = hmc$refresh,
#         control = list(
#           adapt_delta = hmc$adapt_delta, 
#           max_treedepth = hmc$max_treedepth
#         ),
#         file = str_glue("output/brm-2-se-{cur_group()}")
#       )
#     )
#   )
```

```{r}
f2d <- fits_2 %>% 
  mutate(
    out = 
      map(
        fit, 
        ~as_draws_df(., variable = "^b_|^sd_", regex = TRUE) %>% 
          summarise_draws(samples = length, default_convergence_measures())
      )
  ) %>% 
  select(-fit) %>% 
  unnest(out)
f2d %>%  
  kbl(
    caption = "HMC Diagnostics, Model 2c",
    digits = c(0,0,0,2,0,0)
  ) %>% 
  kable_paper(html_font = Font, full_width = FALSE) %>% 
  column_spec(
    4, 
    color = "white",
    background = spec_color(
      f2d$rhat, 
      direction = 1, 
      option = "C", 
      end = .7
    )
  ) %>% 
  scroll_box(height = "600px", width = "800px")
```

```{r}
tmp <- fits_2 %>% 
  mutate(
    out = map(
      fit, 
      ~pp_check(
        .x, 
        ndraws = 10, 
        type = "dens_overlay_grouped",
        group = "region",
        resp = "val", 
        newdata = .x$data
      ) + 
        facet_wrap("group", nrow = 1, scales = "free_y")
    )
  )
wrap_plots(tmp$out, nrow = 3) &
  scale_x_continuous(
    breaks = pretty_breaks(2),
    labels = function(x) percent(x/100, 1)
  )
```

### Summary

We then display the model's population-level parameters

```{r}
tmp <- fits_2 %>% 
  mutate(
    out = map(
      fit, 
      ~as_draws_df(.x, variable = "^b_", regex = TRUE) %>% 
        summarise_draws(
          mean, quantile, rhat, 
          .args = list(probs = c(.025, .975))
        )
    )
  ) %>% 
  select(-fit) %>% 
  unnest(out) %>% 
  mutate(across(where(is.numeric), ~round(., 2))) %>% 
  mutate(
    result = str_glue(
      "b = {mean} ({`2.5%`}, {`97.5%`})"
    )
  ) %>% 
  rowwise() %>% 
  select(-c(mean, `2.5%`, `97.5%`, rhat)) %>% 
  pivot_wider(names_from = cause, values_from = result)
tmp %>%   
  kbl() %>% 
  kable_paper(html_font = Font) 
```

Coefficients

```{r}
tmp <- fits_2 %>% 
  mutate(
    out = map(
      fit, 
      ~emtrends(.x, var = "i1_cw", by = c("sex", "age"))
    )
  ) %>% 
  select(-fit)
tmp %>% 
  mutate(out = map(out, ~as.data.frame(out))) %>% 
  unnest(out) %>% 
  select(cause, sex, age, i1_cw.trend, lower.HPD, upper.HPD) %>% 
  kbl(digits = 3) %>% 
  kable_minimal(full_width = FALSE)
```

## Model 2 for fixed and mobile

```{r}
bf_2_f <- bf(
  val ~ 
    year * sex + year * age + f1_cw * sex + f1_cw * age + f1_cb + 
    (year * sex + year * age + f1_cw * sex + f1_cw * age | country) +
    (year * sex + year * age + f1_cw * sex + f1_cw * age | region),
  family = gaussian()
)

fits_2_f <- dat %>%
  group_by(cause) %>% 
  summarise(
    fit = list(
      brm(
        formula = bf_2_f, 
        data = cur_data_all(),
        chains = hmc$chains, 
        cores = hmc$cores, 
        threads = hmc$threads,
        iter = hmc$iter, 
        warmup = hmc$warmup, 
        refresh = hmc$refresh,
        control = list(
          adapt_delta = hmc$adapt_delta, 
          max_treedepth = hmc$max_treedepth
        ),
        file = str_glue("output/brm-2-fixed-{cur_group()}")
      )
    )
  )

bf_2_m <- bf(
  val ~ 
    year * sex + year * age + m1_cw * sex + m1_cw * age + m1_cb + 
    (year * sex + year * age + m1_cw * sex + m1_cw * age | country) +
    (year * sex + year * age + m1_cw * sex + m1_cw * age | region),
  family = gaussian()
)

fits_2_m <- dat %>%
  group_by(cause) %>% 
  summarise(
    fit = list(
      brm(
        formula = bf_2_m, 
        data = cur_data_all(),
        chains = hmc$chains, 
        cores = hmc$cores, 
        threads = hmc$threads,
        iter = hmc$iter, 
        warmup = hmc$warmup, 
        refresh = hmc$refresh,
        control = list(
          adapt_delta = hmc$adapt_delta, 
          max_treedepth = hmc$max_treedepth
        ),
        file = str_glue("output/brm-2-mobile-{cur_group()}")
      )
    )
  )
```

>>>>

# MCMC settings
ncores <- min(parallel::detectCores(), 12)
nchains <- 4
options(brms.backend = "cmdstanr")
hmc <- list(
  chains = nchains,
  cores = nchains,
  threads = ncores %/% nchains,
  iter = 1000, 
  warmup = 500,
  refresh = 50,
  adapt_delta = .90, 
  max_treedepth = 10
)
```

@@

- Drop countries with no values
- Transform edge values of internet variable to fit beta
- center year on 2010 (roughly in the middle)
- Intercept and simple effects at 2010
- 1 unit: year
- create 1-year lagged internet adoption variables
- Intercepts are country/grand means
- 1 unit: 100%
- Contrast code factors

```{r}
countries_no_internet <- itu %>% 
  group_by(country) %>% 
  summarise(s = sum(internet, na.rm = TRUE)) %>% 
  filter(s == 0)

itu <- itu %>% 
  filter(!(country %in% countries_no_internet$country))

filter(itu, internet %in% c(0, 1))

itu <- itu %>% 
  mutate(internet = if_else(internet == 0, .0000001, internet)) %>% 
  mutate(internet = if_else(internet == 1, .9999999, internet))

@@

```{r}
# Indicate unique ITU values for MV model
dat <- dat %>% 
  # Unique values exist for each country-year, but must be given for each
  # outcome
  group_by(region, country, year, cause) %>% 
  mutate(itu = 1:n() == 1) %>% 
  ungroup()
```

@@

## Model equations

@@

@@

bf_1_se <- bf(
  bf(val | resp_se(se, sigma = TRUE) ~ 
       year * sex + year * age +
       (year * sex + year * age |c| country) +
       (year * sex + year * age |r| region),
     sigma ~ 1
  ),
  family = gaussian(),
  unused = ~ cause
)

bf_2 <- bf(
  val ~ 
    year * sex + year * age + i1_cw * sex + i1_cw * age + i1_cb + 
    (year * sex + year * age + i1_cw * sex + i1_cw * age | country) +
    (year * sex + year * age + i1_cw * sex + i1_cw * age | region),
  family = gaussian(),
  unused = ~ cause
)

bf_2_f <- bf(
  val ~ 
    year * sex + year * age + f1_cw * sex + f1_cw * age + f1_cb + 
    (year * sex + year * age + f1_cw * sex + f1_cw * age | country) +
    (year * sex + year * age + f1_cw * sex + f1_cw * age | region),
  family = gaussian(), 
  unused = ~ cause
)

bf_2_m <- bf(
  val ~ 
    year * sex + year * age + m1_cw * sex + m1_cw * age + m1_cb + 
    (year * sex + year * age + m1_cw * sex + m1_cw * age | country) +
    (year * sex + year * age + m1_cw * sex + m1_cw * age | region),
  family = gaussian(),
  unused = ~ cause
)
```

## Fit all models

```{r}
models <- tibble(
  model = c("1", "1-se", "2", "2-se", "2-f", "2-m"),
  formula = list(
    bf_itu + bf_1 + set_rescor(FALSE),
    bf_itu + bf_1_se + set_rescor(FALSE),
    bf_2,
    bf_2_se,
    bf_2_f,
    bf_2_m
  )
)
fits <- dat %>% 
  # Could nest but need this to include group key in nested tibble
  group_by(cause) %>% 
  summarise(data = list(cur_data_all())) %>% 
  crossing(models)

# Don't estimate se models for selfharm--doesn't converge
fits <- fits %>% 
  filter(!(cause=="Selfharm" & str_detect(model, "se")))

# Estimate all models
fits <- fits %>% 
  mutate(
    fit = pmap(
      list(cause=cause, data=data, model=model, formula=formula),
      function(cause, data, model, formula)
        brm(
          formula = formula,
          data = data,
          chains = hmc$chains,
          cores = hmc$cores,
          threads = hmc$threads,
          iter = hmc$iter,
          warmup = hmc$warmup,
          refresh = hmc$refresh,
          control = list(
            adapt_delta = hmc$adapt_delta,
            max_treedepth = hmc$max_treedepth
          ),
          file = str_glue("output/brm-{model}-{cause}")
        )
    )
  )
```

@@

```{r}
diag_1 <- fits %>% 
  filter(model == "1") %>% 
  mutate(
    out = 
      map(
        fit, 
        ~as_draws_df(., variable = "^b_|^sd_|^cor_", regex = TRUE) %>% 
          summarise_draws(samples = length, default_convergence_measures())
      )
  ) %>% 
  select(-fit) %>% 
  unnest(out)
diag_1 %>%  
  kbl(
    caption = "HMC Diagnostics, Model 1",
    digits = c(0,0,0,2,0,0)
  ) %>% 
  kable_paper(html_font = Font) %>% 
  column_spec(
    4, 
    color = "white",
    background = spec_color(
      diag_1$rhat, 
      direction = 1, 
      option = "C", 
      end = .7
    )
  )
```

```{r}
ppcheck_1 <- fits %>% 
  filter(model == "1")
  mutate(
    out = map(
      fit, 
      ~pp_check(
        .x, 
        ndraws = 10, 
        type = "dens_overlay_grouped",
        group = "region",
        resp = "val", 
        newdata = .x$data
      ) + 
        facet_wrap("group", nrow = 1, scales = "free")
    )
  )
wrap_plots(ppcheck_1$out, nrow = 6) &
  scale_x_continuous(
    breaks = pretty_breaks(2),
    labels = function(x) percent(x/100)
  ) &
  theme(legend.position = "none")
```
